---
title: "Mediocrity"
author: "Einar Holsbø"
date: "2 January 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="..")
```
```{r, include=F}
library(ProjectTemplate); load.project()  # load project
library(ggthemes)
theme_set(theme_tufte())
```

Do crime rates regress toward mediocrity? If they did, it would make 100% sense 
to do shrinkage.

## Regresession Toward Mediocrity in Hereditary Violence

```{r violence, warning=F, message=F}
violence <- norwegian_crime %>% filter(crime_type == "Voldskriminalitet") %>%
            select(-crime_type) %>% transmute(place, year, rate=reports/population)

years <- violence %>% spread(year, rate)


# maybe this can be done prettier......
p_1 <- ggplot(years) + aes(x=`2010`, y=`2011`) + geom_point(color="grey50") +
  stat_smooth(method = "lm", se = TRUE) + geom_abline(slope=1, intercept=0) +
  geom_abline(slope=0, intercept=mean(violence$rate)) 
p_2 <- ggplot(years) + aes(x=`2011`, y=`2012`) + geom_point(color="grey50") +
  stat_smooth(method = "lm", se = TRUE) + geom_abline(slope=1, intercept=0) +
  geom_abline(slope=0, intercept=mean(violence$rate)) 
p_3 <- ggplot(years) + aes(x=`2012`, y=`2013`) + geom_point(color="grey50") +
  stat_smooth(method = "lm", se = TRUE) + geom_abline(slope=1, intercept=0) +
  geom_abline(slope=0, intercept=mean(violence$rate)) 

library(gridExtra)
grid.arrange(p_1, p_2, p_3, ncol=3)
```

The above plots show crime rate comparisons for different pairs of years. The
horizontal line is the mean crime rate through all years, the solid dark line is
$x=y$ (indicating no change), and the blue line is the least squares regression
line in these data. As we can see, the regression line does indeed tip toward
the mean. This was true for the handful of pairwise comparisons I did by hand when
writing this.

## What about smoothing??
So let's do smoothing for 2012+2013 and compare.
```{r smooth, warning=F, message=F}
mu <- mean(years$`2012`, na.rm = T)
sigsq <- var(years$`2012`, na.rm = T)
alpha_p <- ((1 - mu) / sigsq - 1 / mu) * mu ^ 2
beta_p <- alpha_p * (1 / mu - 1)

twelve <- norwegian_crime %>% 
  filter(year==2012, crime_type=="Voldskriminalitet") %>%
  select(place, reports, population) %>%
  transmute(place, `2012s`=(reports+alpha_p)/(population+alpha_p+beta_p))
years %<>% left_join(twelve)

mu <- mean(years$`2013`, na.rm = T)
sigsq <- var(years$`2013`, na.rm = T)
alpha_p <- ((1 - mu) / sigsq - 1 / mu) * mu ^ 2
beta_p <- alpha_p * (1 / mu - 1)

thirteen <- norwegian_crime %>% 
  filter(year==2013, crime_type=="Voldskriminalitet") %>%
  select(place, reports, population) %>%
  transmute(place, `2013s`=(reports+alpha_p)/(population+alpha_p+beta_p))
years %<>% left_join(thirteen)

lim <- c(0.001, 0.015)
p_1 <- ggplot(years) + aes(x=`2012`, y=`2013`) + geom_point(color="grey50") +
  stat_smooth(method = "lm", se = TRUE) + geom_abline(slope=1, intercept=0) +
  geom_abline(slope=0, intercept=mean(violence$rate)) +
  coord_cartesian(xlim = lim, ylim=lim)
p_2 <- ggplot(years) + aes(x=`2012s`, y=`2013s`) + geom_point(color="grey50") +
  stat_smooth(method = "lm", se = TRUE) + geom_abline(slope=1, intercept=0) +
  geom_abline(slope=0, intercept=mean(violence$rate)) +
  coord_cartesian(xlim = lim, ylim=lim)

grid.arrange(p_1, p_2, ncol=2)
```

There seems to be the same average amount of regression to mediocrity using 
smoothing vs. not using smooting. This might make sense, i don't think this is a
phenomenon we can magic away by clever modelling. We do however get rid of many 
high-leverage/outlying points, it seems. There is probably some kind of
decreased MSE argument hiding here.

Anyway if we plot the two regression lines in the same plot (black is
non-smoothed), they're pretty much indistiguishable:

```{r regr, warning=F}
ggplot(years) + 
  stat_smooth(aes(`2012`, `2013`), method="lm", color="black", se = F) +
  stat_smooth(aes(`2012s`, `2013s`), method="lm", color="blue") 
```
