---
title: "Simulation"
author: "Einar Holsbø"
date: "5 October 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="..")
```
```{r, include=F}
library(ProjectTemplate); load.project()  # load project
```
# Q: should one simulate from a single common crime rate or several different ones?
If we're going to simulate a year of crime rates per person (pp.) it's sensible 
to look at how the crime rates pp. are distributed for each year. We have
data from 2008 to 2014. In the below plot the coloms are some of the years we have data for,
top row is observed crime rates that year $\{\lambda_i\}$, 
middle row is to simulate from the observed crime rates by 
$\hat{cpp_i} = c_i/p_i$, with $c_i \sim Pois(\lambda_i)$, and the bottom row is
the same as the middle but with the common crime rate $c_i \sim Pois(median(\{\lambda_i\}))$
```{r}
years <- as.character(2011:2014)
total_c <- total_crime() 
total_c %<>% mutate(rate=reports/population)

par(mfcol=c(3,length(years)))
for (y in years) {
  total_c %>% filter(year==y) %>% select(rate) %>% t %>% density %>% plot(main="")
  total_c %>% filter(year==y) %>% 
              mutate(rate=simulate_crimes(population, rate)/population) %>%
              select(rate) %>% t %>% density %>% plot(main="")
  total_c %>% filter(year==y) %>% 
              mutate(rate=simulate_crimes(population, median(rate))/population) %>%
              select(rate) %>% t %>% density %>% plot(main="")
}
```

Clearly it's no good to simulate from a single common crime rate. We'll use the
observed crime rates. To ameliorate the fact that there's probably a lot of variance
in the observed crime rates from a given year, we'll use the mean rate from all years
as ground truth in our simulations.

The "true" populations & crime rates:

```{r}
towns <- mean_crimes()

plot(towns, pch=20)
```

## Evaluation of smoothing
Let's draw from the simulated truth thousands of times and see where this lands
us in terms of errors.
```{r}
errors <- raply(10000, {
  towns %<>% mutate(crimes=simulate_crimes(towns$population, towns$cpp), raw_est=crimes/population)
  
  mu <- mean(towns$raw_est)
  sigsq <- var(towns$raw_est)
  alpha_p <- ((1 - mu) / sigsq - 1 / mu) * mu ^ 2
  beta_p <- alpha_p * (1 / mu - 1)
  
  towns %<>% mutate(smoothed_est=(crimes+alpha_p)/(population+alpha_p+beta_p))
  
  error_raw <- mean(abs(towns$raw_est - towns$cpp))
  error_smooth <- mean(abs(towns$smoothed_est - towns$cpp))
  c(raw=error_raw, smooth=error_smooth)
})

plot(density(errors[,2]), col="red")
abline(v=mean(errors[, 2]), col="red")

lines(density(errors[, 1]))
abline(v=mean(errors[, 1]))
```
Smoothed estimates in red, there is evidence that, on average, there's 
a lower error for the smoothing. 

# Evaluation of updating estimate year by year
TODO:

* Spearman rank correlation for raw estimates vs adjusted.
* Local FDR-based rankings?
* Simulate from ground truth n times to create n virtual years
* Estimate prior either from all years or from first year
* Update estimates year by year by using the posterior from one year as prior for the next

